### Loading and preprocessing the data
```{r}
data <- read.csv('activity.csv')
head(data)
summary(data)
```
### What is mean total number of steps taken per day?


```{r message=FALSE}
mean_median_steps <- function(data)
{
        library(dplyr)
        by_data <- group_by(data,date)
        steps_each_day <- summarise(by_data,sum(steps,na.rm = TRUE))
        steps_each_day
        colnames(steps_each_day)[2]="total_steps"
        
        library(ggplot2)
        print(qplot(total_steps, data=steps_each_day))
        mean_step <- mean(steps_each_day$total_steps)
        median_step <- median(steps_each_day$total_steps)
        list(mean_step=mean_step,median_step=median_step)
}
mean_median <- mean_median_steps(data)
```
   
Mean of the total of steps taken per day is **`r mean_median["mean_step"]`**  
Median of the total of steps taken per day is **`r mean_median["median_step"]`**

### What is the average daily activity pattern?

```{r}
plot_steps_interval <- function(data,facets=NA)
{
        avg_daily_activity <- summarise(group_by(data, interval),mean(steps,na.rm=TRUE))
        colnames(avg_daily_activity)[2] <- "steps"
        max_step <- max(avg_daily_activity$steps)
        max_interval <- as.numeric(select(filter(avg_daily_activity,steps==max(steps)),
                                          interval))
        if(is.na(facets))
        {
                print(ggplot(avg_daily_activity,aes(x=interval,y=steps))
                +geom_line(col="blue")+geom_vline(xintercept=max_interval,
                linetype="dotted", col="red")+geom_hline(yintercept=max_step,
                linetype="dotted",col="red"))
        }
        else
        {
                print(ggplot(avg_daily_activity,aes(x=interval,y=steps))
                +geom_line(col="blue")+facet_grid(~day))
        }
        max_interval
}
max_interval <- plot_steps_interval(data)
```
   
The 5-minute interval, on average across all the days in the dataset, which contains the maximum number of steps is **`r max_interval`**

### Imputing missing values
```{r}
missing_indices <- is.na(data[,"steps"])
total_row_with_NA <- sum(missing_indices)

avg_daily_activity <- summarise(group_by(data, interval),mean(steps,na.rm=TRUE))
colnames(avg_daily_activity)[2] <- "steps"
newdata <- data
newdata[missing_indices,"steps"]<- avg_daily_activity[match(data[missing_indices,"interval"],avg_daily_activity$interval),"steps"]

mean_median <- mean_median_steps(newdata)
max_interval <- plot_steps_interval(newdata)
```   
   
The total number of rows with NA is `r total_row_with_NA`   
   
The missing values are replaced with mean of 5-minute interval.   
In new data,   
        Mean of the total of steps taken per day is **`r mean_median$mean_step`**     
        Median of the total of steps taken per day is **`r mean_median$median_step`**   
        The 5-minute interval, on average across all the days in the dataset, which contains the maximum number of steps is **`r max_interval`**   

The *mean* and *median* value of the total steps taken per day has **increased**.
   
### Differences in activity patterns between weekdays and weekends

```{r}
newdata$day <- as.factor(weekdays(as.Date(newdata$date)) %in% c("Sunday","Saturday"))
levels(newdata$day) <- c("weekday","weekend")

avg_daily_activity <- summarise(group_by(newdata, interval,day),mean(steps,na.rm=TRUE))
colnames(avg_daily_activity)[3] <- "steps"
print(ggplot(avg_daily_activity,aes(x=interval,y=steps))+
              geom_line(col="blue")+facet_grid(day~.))
```